### 2019-04-29 report/papers.tex
on continue la review

>>>>

\paragraph{Using a heat kernel in the entropy-regularization}
We recall what is the entropy-regularized Wasserstein distance between $\mu_0$ and $\mu_1$ on the domain $M$:
$$ W_{2, \gamma}^2(\mu_0, \mu_1) = \inf_{\pi \in \Pi} \left[ \int \int_{M \times M} d(x, y)^2 \pi(x, y) dx dy \, - \, \gamma H(\pi) \right] = \gamma \left[ 1 + \min_{\pi \in \Pi} \text{KL}(\pi | \mathcal{K}_\gamma) \right]$$
Where:
$$ \Pi = \left\{ \pi \in \text{Prob}(M \times M) \, | \, \pi(\cdot, M) = \mu_0, \pi(M, \cdot) = \mu_1 \right\} $$
$$ H(\pi) = - \int \int_{M \times M} \pi(x, y) \ln \pi(x, y) dx dy $$
$$ \mathcal{K}_\gamma = e^{-d(x, y)^2 / \gamma} $$
$$ \text{KL}(\pi | \mathcal{K}) = \int \int_{M \times M} \pi(x, y) \left[ \ln \dfrac{\pi(x, y)}{\mathcal{K}(x, y)} - 1 \right]\ln  dx dy $$
This is a strictly convex problem thanks to the entropy. The idea of the paper is to use heat kernel because is some cases, the solution of the diffusion equation can be computed in a effective way. We denote by $\mathcal{H}_t(x, y)$ the diffusion between $x$ and $y$ after a time $t$ in the heat kernel. We have:
$$ \mathcal{K}_\gamma \approx \mathcal{H}_{\gamma / 2}(x, y) $$
We won't store $\mathcal{H}$ because it has a space complexity in $\mathcal{O}(n^2)$ and we want a lower complexity. But we generally know a way to apply $\mathcal{H}$ to a vector. In the case of images we apply Gaussian convolution with $\sigma^2 = \gamma$. \\
In the case of triangle meshes we associate a weight to faces proportional to their area. We denote by $a$ the vector of these weights (In images of size $n \times m$ we set $a = 1 / (nm)$). Then we denote by $L$ the cotangent Laplacian and by $D_a$ the diagonal matrix with diagonal $a$. By discretizing the heat equation, we obtain:
$$ w = \mathcal{H}_t(v) \Leftrightarrow \left( D_a + t L \right) w = v $$
The linear system can be solved efficiently by pre-computing a sparse Cholesky factorization.

\paragraph{Algorithm}
We some computations we arrive at:

\vspace{3mm}
\begin{algorithm}[H]
	\caption{\textsc{Convolutional-Sinkhorn}($\mu_0, \mu_1, H_t, a$)}
	$v, w \gets 1$
	\Repeat{convergence of $u, v$}{
		$v \gets \mu_0 \oslash H_t(a \otimes w)$ \;
		$v \gets \mu_1 \oslash H_t(a \otimes wv)$ \;
	}
	\Return $2t a^\trans \left[ \left( \mu_0 \otimes \ln v \right) + \left( \mu_1 \otimes \ln w \right) \right]$
\end{algorithm}
\vspace{3mm}

Where $\oslash$ and $\otimes$ denote element-wise operations. As Sinkhorn algorithm we obtain a simple iterative algorithm where, this time, steps can be computed in $\mathcal{O}(n \log n)$ in many cases, where $n$ is the size of the domain $M$.

\paragraph{Barycenter}
An algorithm for computing barycenters is also provided. We won't enter into details but we give their algorithm for computing the barycenter of distributions $\mu_i$ associated with weights $\alpha_i$:

\vspace{3mm}
\begin{algorithm}[H]
	\caption{\textsc{Convolutional-Barycenter}($\mu_0, \mu_1, H_t, a$)}
	$v, w \gets 1$
	\Repeat{convergence of $u, v$}{
		$v \gets \mu_0 \oslash H_t(a \otimes w)$ \;
		$v \gets \mu_1 \oslash H_t(a \otimes wv)$ \;
	}
	\Return $2t a^\trans \left[ \left( \mu_0 \otimes \ln v \right) + \left( \mu_1 \otimes \ln w \right) \right]$
\end{algorithm}
\vspace{3mm}
